---
title: "compmus"
author: "popke"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    storyboard: true
    theme: lumen
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
source("compmus.R")
get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}  


library(rjson)
library(tidymodels)
library(ggdendro)
library(heatmaply)
library(patchwork)
library(ggthemes)

compmus2025 <- read_csv("compmus2025.csv")

cluster_juice <-
  recipe(
    filename ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(compmus2025) |>
  juice() |>
  column_to_rownames("filename")

compmus2025_filtered <- 
  compmus2025 |> filter(!is.na(ai)) |> 
  mutate(ai = factor(if_else(ai, "AI", "Non-AI")))

classification_recipe <-
  recipe(
    ai ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025_filtered
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors())      # Converts to z-scores.
  # step_range(all_predictors())    # Sets range to [0, 1].

compmus_cv <- compmus2025_filtered |> vfold_cv(5)

#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )


```

### AI analysis class of 25

This dashboard will be a demonstration of correlation between multiple musical terms like tempo, timbre, loudness and other features. 
Comparing and analyzing music gives us a better insight in how songs are constructed and how we can interpret them. 
For this course we use a data set that consists of songs that were created for this course with AI, composed or already existing songs. The class corpus is sufficiently well distributed and we therefore can compare AI tracks with non-AI tracks and analyse their characteristics against each other. This course has given us a deeper understanding of the metrics of music and gave us metrics to analyze the songs. Do AI tracks lack timbre and if yes, in what kind of way? and more importantly: "Can we recognize the difference between AI and non-AI tracks?" 

I've chosen my tracks because one of them is made by artist I work with and am a huge fan of and the other is made by a friend who also produces music. Both tracks are electronic music tracks, in the genre Techno. They are quiet similar genre wise, but are different in comparison. 
Track 1 is called: "echoes in the abyss" composed by Neither nor. 
Track 2 is called: "paper battles" composed by Parabool. 
Track 1 is a more melodic track with soft lyrics, it gives you more the feeling of flowing trough the track like a river. 
Track 2 is more hectic, with a faster BPM. The sub tones in the track feel more chaotic, but also uses vocals. A bit darker than song 1. The chords stay closer together. But make use of build ups and releases. I think my tracks match the class corpus, I came across a lot of instrumental music with a lot melodies in it. I think it's an interesting mix of music. 

Our musical corpus is relatively good devided in AI and non-AI tracks. In the next few chapters we will walk trought the music and analyse the corpus. Discussing our findings as we go along. 
Enjoy!


***

#### SONG 1: Neither Nor - Echoes in the Abyss
<audio controls>
    <source src="audio/popke-s-1.mp3" type="audio/mp3">
    Your browser does not support the audio element.
</audio>

---

#### SONG 2: Parabool - Paper Battles 
<audio controls>
    <source src="audio/popke-s-2.mp3" type="audio/mp3">
    Your browser does not support the audio element.
</audio>



### analysis of class corpus: dendogram

```{r, analysis of class corpus}

compmus_dist <- dist(cluster_juice, method = "euclidean")

compmus_dist |> 
  hclust(method = "average") |> # Try single, average, and complete.
  dendro_data() |>
  ggdendrogram()

```

***

The dendogram is created by measuring the average distance between the tracks in the corpus. On the left side we can see a few outliers. other than that we can see three big clusters. The biggest one on the right side, and two smaller ones on the more left side. 
The dendogram gives a good visualisation about the musical similarity in our class corpus. It shows us where the majority of the songs are categorised and which songs are the most farout outliers. 

### analysis of class corpus: heatmap

```{r, heatmap}
heatmaply(
  cluster_juice,
  hclustfun = hclust,
  hclust_method = "average",  # Change for single, average, or complete linkage.
  dist_method = "euclidean"
)
```

***

In this heatmap we compare "valence", "arousal", "danceability", "instrumentalness", "tempo". On the right side and top we can see the clusters created by comparing all these characteristics. 

### analysis of class corpus: k nearest neighbours
```{r}

knn_model <-
  nearest_neighbor(neighbors = 1) |>
  set_mode("classification") |> 
  set_engine("kknn")
classification_knn <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(knn_model) |> 
  fit_resamples(compmus_cv, control = control_resamples(save_pred = TRUE))

classification_knn |> get_conf_mat()

classification_knn |> get_pr()

classification_knn |> get_conf_mat() |> autoplot(type = "heatmap")




```

***

With the method of K-nearest neighbours we can take a deeper look into the specifics of the clustering algorithm to see what is going on. 
If we look at the precision and recall scores we can see that these are not that high. 

precision AI tracks: 0.648
recall AI tracks: 0.714
precision non-AI tracks: 0.611
recall non-AI tracks: 0.537

This means the model is better in predicting AI tracks than in non AI tracks, which is logical considering the AI tracks are more similar than the non-AI tracks. And therefore loosing a lot when trying to find non-AI tracks.  


### analysis of class corpus: AI or no AI

```{r}

compmus_cv <- compmus2025_filtered |> vfold_cv(5)
compmus2025_filtered |>
  ggplot(aes(x = valence, y = arousal, colour = ai, size = tempo)) +
  geom_point(alpha = 0.8) +
  scale_color_viridis_d() +
  labs(
    x = "Valence",
    y = "Arousal",
    size = "Tempo",
    colour = "AI"
  )
```

***
We see here a bubble scatter plot of the whole class corpus where valence is placed on the x-axes and arousal on the y axis. We see that the spread of valence and arousal is quiet similar. The AI songs are more densely distributed in the middle of the graph, but also have a few outliers withe higher arousal and valence scores. This can be because AI songs are more energetic or uplifting. The tempo differences in AI generated songs are fairly similar where the non-AI songs fluctuate a lot. 
When we compare AI-generated music with human-made music we can see that human-made music still contains a lot mor diversity than the AI tracks show in this graph. 

### spectral & energy novelty

```{r, energy_novelty}
  data1 <-compmus_energy_novelty("features/popke-s-1.json") 
  fig1 <- (ggplot(data1, aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Energy Novelty song1", x = "Time (s)", y = "Energy Novelty"))


  data2 <- compmus_energy_novelty("features/popke-s-2.json")
  fig2 <- (ggplot(data2, aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Energy Novelty song 2",x = "Time (s)", y = "Energy Novelty"))

  data3 <- compmus_spectral_novelty("features/popke-s-1.json")
  fig3 <- (ggplot(data3, aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Spectral Novelty song 1", x = "Time (s)", y = "Spectral Novelty")) 
  
  data4 <- compmus_spectral_novelty("features/popke-s-2.json")
  fig4 <- (ggplot(data4, aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Spectral Novelty song 2", x = "Time (s)", y = "Spectral Novelty"))

  
  fig1 / fig2 / fig3 / fig4

```

***

Spectral & Energy based novelty graphs are used when we want to find parts in a song or music pieces that show a burst of novel or energetic music. 
Where the energy novelty shows information about the changes in loudness or transitions in the music. The spectral energy gives us more information about instrumental shifts, harmonic shifts or changes in timbre. 
If we look at song 1 we see peaks in the energy novelty at around 10, 90 and 150 seconds. When we listen to the track we first hear the intro and after about 10 seconds the base starts playing. At 50 seconds we hear a few soft vocals, which is also explained by the little bump at +50 seconds. After about 90 seconds the female voice comes in play. At around 150 seconds we see a few peaks close together in the graph at this point, the voice sings acapella and goes from a very high pitch to a lower pitch where after the music comes back in. The spectral novelty shows a lot of peaks throughout the song, it's lower when the beat stops and then comes back up again. It gives us more global information about the harmony, while the energy novelty can point to parts in the song where something new is happening. 

For song 2 this happens at around 50 and 140 seconds.At 50 seconds after a break in the music the beat comes back what is shown in the energy novelty. At around 140 seconds the beat stops and dark vocals come in. Because the song uses a more repetitive structure, the spectral novelty graph doesn't show a lot of changes throughout the track. 

We can see energy novelty and spectral novelty can give us a lot of different information about what is happening in a track. When the music makes use of a more complicated structural harmonic section this would be something where a spectral novelty graph could be more efficient than when analysing repetitive structures. 



### Tempograms

```{r, tempograms}
  data1 <- compmus_tempogram("features/popke-s-1.json", window_size = 4, hop_size = 2, cyclic = FALSE)
  fig1 <- ggplot(data1, aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

  data2 <- compmus_tempogram("features/popke-s-2.json", window_size = 4, hop_size = 2, cyclic = FALSE)
  fig2 <- ggplot(data2, aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
  
  fig <- subplot(fig1, fig2, nrows=2)%>%
  layout(
    title = "tempograms",
    xaxis = list(title = "Time"),
    yaxis = list(title = "Temo (BPM)")
  )
  fig
```


***
Here we see the tempo gram plots for both songs. I first started out with a window size of 32 and hop size of 4. Where-after i got only really tight lines in my tempo grams. I adjusted the window size first to 8 and kept the hop size the same. The fluctuations in the tempo gram became already clearer. With lowering the window size you can clearly see more fluctuations and details in the tempo gram. We can see a clear tempo at ±135 BPM, which is also the tempo of the song, then we see 3 subharmonics at ±275, ±405, ±524. These are all doublings of the base tempo and therefore picked up in the tempograph. 


### cyclic tempograms

```{r, cyclic tempograms}
  data1 <- compmus_tempogram("features/popke-s-1.json", window_size = 4, hop_size = 2, cyclic = TRUE)
  fig1 <- ggplot(data1, aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

  data2 <- compmus_tempogram("features/popke-s-2.json", window_size = 4, hop_size = 2, cyclic = TRUE)
  fig2 <- ggplot(data2, aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
  
  fig <- subplot(fig1, fig2, nrows=2)%>%
  layout(
    title = "tempograms",
    xaxis = list(title = "Time"),
    yaxis = list(title = "Tempo (BPM)")
  )
  fig
```

***

In the cyclic tempo gram, we have a more visible sight of the actual tempo of the songs. Which is arond 137 for both of them. 


### keychrams and Chordchrams

```{r, Keycharams & Choardchrams}

  data1 <- compmus_chroma("features/popke-s-1.json", norm = "identity")%>%
  compmus_match_pitch_templates(
    key_templates,         
    norm = "identity",       
    distance = "cosine")
  
  fig1 <- ggplot(data1, aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(title = "keygram song 1", y = "Keys", fill = NULL) + theme_classic() + theme(plot.title = element_text(size=8), axis.title.y = element_text(size=8))

  data2 <- compmus_chroma("features/popke-s-1.json", norm = "identity")%>%
    compmus_match_pitch_templates(
    chord_templates,         
    norm = "identity",       
    distance = "cosine"   
  ) 
  
  fig2 <- ggplot(data2, aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +           
  labs(title = "chordgram song 1", y = "Chords", fill = NULL) + theme_classic() + theme(plot.title = element_text(size=8),  axis.title.y = element_text(size=8))

    data3 <- compmus_chroma("features/popke-s-2.json", norm = "identity")%>%
  compmus_match_pitch_templates(
    key_templates,         
    norm = "identity",       
    distance = "cosine")
  
  fig3 <- ggplot(data3, aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               
  labs(title = "keygram song 2", y = "Keys", fill = NULL) + theme_classic() + theme(plot.title = element_text(size=8), axis.title.y = element_text(size=8))

  data4 <- compmus_chroma("features/popke-s-2.json", norm = "identity")%>%
    compmus_match_pitch_templates(
    chord_templates,         
    norm = "identity",       
    distance = "cosine"   
  ) 
  fig4 <- ggplot(data4, aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               
  labs(title = "chordgram song 2", x = "Time (s)", y = "Chords", fill = NULL) + theme_classic() + theme(plot.title = element_text(size=8), axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
  
fig1 / fig2 / fig3 / fig4 
  
```

***
Here we see a representation of key grams and chord grams for both songs. Key grams are used to show the estimated musical key used in a track over time. Chordgrams on the other hand show the estimated chords. 

When we look at the keygrams from both songs, we see that song 1 has more harmonic variety, while song 2 maintains more tonal consistency. Both songs show a big trasnsition as well. 
When we look at the chordgrams, we see the chords in song 1 show frequent chord changes, which indicates an active harmonic rhythm. Song 2 also shows frequent chord changes but these are more evenly distributed. 
If we look at the chord and key grams, we can say that song 1 shows more harmonics while song 2 is more rhythmically consistent. Which we can hear in the songs posted on the first page of this flexdashboard. 

### Chromagrams song 1 & 2

```{r, Chromagrams}
  data1 <- compmus_chroma("features/popke-s-1.json", norm = "identity")      
  fig1 <- ggplot(data1, aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()  

  data2 <- compmus_chroma("features/popke-s-2.json", norm = "identity")
  fig2 <- ggplot(data2, aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +              
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()  
  
  fig <- subplot(fig1, fig2)
  fig
```

***

A chromagram is a time versus pitch class matrix, which is often used in Music Information Retrieval (MIR). It shows how much energy there is present for each of the 12 pitch classes. While a spectogram shows all frequencies, a chromagram visualises the frequencies for each octave in a single representation. 
When looking at the left chromagram, we can see that it appears a bit denser with a few more bright patches which suggest the use of louder notes or more harmonic complexity. The right chromagram has a more evenly divided distribution. We can see a lot of energy in the tonal use at the E, F and G chords. 
This could indicate that song 2 is played in an F major or D minor scale whereas song 1 could be played in A minor or G major. 
The left chromagram shows a few sections where you can see distinctions in harmonic shifts whereas the right chromagram stays more consistent over time. This might suggest a minimal song or a repetive structure, which we can also hear in the music. 


### cepstograms song 1 and 2 

```{r, Cepstograms}
  data1 <- compmus_mfccs("features/popke-s-1.json", norm = "identity")
  fig1 <- ggplot(data1, aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +              
  labs(x = "Time (s)", y = "Coefficient Number song 1", fill = NULL) +
  theme_classic()                                

  data2 <- compmus_mfccs("features/popke-s-2.json", norm = "identity")
  fig2 <- ggplot(data2, aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +              
  labs(x = "Time (s)", y = "Coefficient Number song 2", fill = NULL) +
  theme_classic()

 fig1 / fig2 

```

***
This image shows the two cepstograms for both songs, they visualize the evolution of Mel-Fequency Cepstral Coefficients (MFCCs) over time. MFCCs are used for speech and audio processing because they capture the spectral characteristics of timbre and phonetic features. 
Both cepstograms share a similar overall structure. The right one shows a more smoothed and uniform distribution, which indicates a more stable spectral envelope. Which fits with the song, where tones are very stable and there are not many fluctuations. 
The left cepstogram shows more variation in the horizontal band on the bottom, in the blue part. This indicates more energy in the lower cepstral coefficients. We can distinguish a difference in spectral energy in both songs. 


### Data applied on test visualisation

```{r,datavisualisation}
aisc2024 <- read_csv("compmus2025.csv")
aisc2024 |>                   # Start with the data
  ggplot(                     # Set up the plot.
    aes(
      x = tempo,
      y = arousal,
      size = instrumentalness,
      colour = danceability
    )
  ) +
  geom_point() +              # Scatter plot.
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
  geom_text(                  # Add text labels from above.
    x = 121,
    y = 4.91,
    label = "Onda Corta - Sud America",
    size = 3,                 # Override size (not loudness here).
    hjust = "left",           # Align left side of label with the point.
    vjust = "center",         # Align vertical centre of label with the point.
    angle = 30                # Rotate the text label
  ) +
  scale_x_continuous(         # Fine-tune the x axis.
    limits = c(50, 200),
    breaks = c(50, 100, 150, 200), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(1, 9),
    breaks = c(1, 5, 9),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # Use the popular viridis colour palette.
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Tempo",
    y = "Arousal",
    colour = "Danceability"
  )
ggplotly()
```


### histogram tempo

```{r, histogram data}

cm2025 <- read_csv("compmus2025.csv")
  ggplot(cm2025, aes(x= tempo)) + geom_histogram()
  ggplotly()
```

***

This is a histogram that gives insight to the tempo variation of our class corpus. The median is at 85 and the mean is around 100. 
I think the analysis isn't really specific, for both my own songs the tempo was 30 bpm lower than i saw in the file. I think the data is not that reliable. 


### correlation tempo and danceability

```{r, correlation tempo/danceability}
cm2025 <- read_csv("compmus2025.csv")
ggplot(cm2025, aes(x = tempo, y = danceability)) +
  geom_point(aes(color = engagingness), alpha = 0.7) +
  theme_minimal() +
  labs(title = "Tempo vs. Danceability", x = "Tempo", y = "Danceability") +
  scale_color_gradient(low = "purple", high = "orange")
ggplotly()
```




