---
title: "compmus"
author: "popke"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    storyboard: true
    theme: lumen
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
source("compmus.R")
get_conf_mat <- function(fit) {
  outcome <- .get_tune_outcome_names(fit)
  fit |> 
    collect_predictions() |> 
    conf_mat(truth = outcome, estimate = .pred_class)
}  

get_pr <- function(fit) {
  fit |> 
    conf_mat_resampled() |> 
    group_by(Prediction) |> mutate(precision = Freq / sum(Freq)) |> 
    group_by(Truth) |> mutate(recall = Freq / sum(Freq)) |> 
    ungroup() |> filter(Prediction == Truth) |> 
    select(class = Prediction, precision, recall)
}  


library(rjson)
library(tidymodels)
library(ggdendro)
library(heatmaply)
library(patchwork)
library(ggthemes)

compmus2025 <- read_csv("compmus2025.csv")

cluster_juice <-
  recipe(
    filename ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors()) |> 
  # step_range(all_predictors()) |> 
  prep(compmus2025) |>
  juice() |>
  column_to_rownames("filename")

compmus2025_filtered <- 
  compmus2025 |> filter(!is.na(ai)) |> 
  mutate(ai = factor(if_else(ai, "AI", "Non-AI")))

classification_recipe <-
  recipe(
    ai ~
      arousal +
      danceability +
      instrumentalness +
      tempo +
      valence,
    data = compmus2025_filtered
  ) |>
  step_center(all_predictors()) |>
  step_scale(all_predictors())      # Converts to z-scores.
  # step_range(all_predictors())    # Sets range to [0, 1].

compmus_cv <- compmus2025_filtered |> vfold_cv(5)

#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )


```

### Welcome

This dashboard will be a demonstration of correlation between multiple musical terms like tempo, timbre, loudness and other features. 
Comparing and analyzing music gives us a better insight in how songs are constructed and how we can interpret them. 
For this course we use a data set that consists of songs that were created for this course with AI, composed or already existing songs. The class corpus is sufficiently well distributed and we therefore can compare AI tracks with non-AI tracks and analyse their characteristics against each other. This course has given us a deeper understanding of the metrics of music and gave us metrics to analyze the songs. Do AI tracks lack timbre and if yes, in what kind of way? and more importantly: "Can we recognize the difference between AI and non-AI tracks?" 

I've chosen my tracks because one of them is made by artist I work with and am a huge fan of and the other is made by a friend who also produces music. Both tracks are electronic music tracks, in the genre Techno. They are quiet similar genre wise, but are different in comparison. 
Track 1 is called: "echoes in the abyss" composed by Neither nor. 
Track 2 is called: "paper battles" composed by Parabool. 
Track 1 is a more melodic track with soft lyrics, it gives you more the feeling of flowing trough the track like a river. 
Track 2 is more hectic, with a faster BPM. The sub tones in the track feel more chaotic, but also uses vocals. A bit darker than song 1. The chords stay closer together. But make use of build ups and releases. I think my tracks match the class corpus, I came across a lot of instrumental music with a lot melodies in it. I think it's an interesting mix of music. 


***

#### SONG 1: Neither Nor - Echoes in the Abyss
<audio controls>
    <source src="audio/popke-s-1.mp3" type="audio/mp3">
    Your browser does not support the audio element.
</audio>

---

#### SONG 2: Parabool - Paper Battles 
<audio controls>
    <source src="audio/popke-s-2.mp3" type="audio/mp3">
    Your browser does not support the audio element.
</audio>



### analysis of class corpus

```{r, analysis of class corpus}

compmus_dist <- dist(cluster_juice, method = "euclidean")

compmus_dist |> 
  hclust(method = "average") |> # Try single, average, and complete.
  dendro_data() |>
  ggdendrogram()

```

***

The dendogram is created by measuring the average distance between the tracks in the corpus. On the left side we can see a few outliers. other than that we can see three big clusters. The biggest one on the right side, and two smaller ones on the more left side. 

### heatmap

```{r, heatmap}
heatmaply(
  cluster_juice,
  hclustfun = hclust,
  hclust_method = "average",  # Change for single, average, or complete linkage.
  dist_method = "euclidean"
)
```

***

In this heatmap we compare "valence", "arousal", "danceability", "instrumentalness", "tempo". On the right side and top we can see the clusters created by comparing all these characteristics. 

### k nearest neighbours
```{r}

knn_model <-
  nearest_neighbor(neighbors = 1) |>
  set_mode("classification") |> 
  set_engine("kknn")
classification_knn <- 
  workflow() |> 
  add_recipe(classification_recipe) |> 
  add_model(knn_model) |> 
  fit_resamples(compmus_cv, control = control_resamples(save_pred = TRUE))

classification_knn |> get_conf_mat()

classification_knn |> get_conf_mat() |> autoplot(type = "mosaic")

classification_knn |> get_conf_mat() |> autoplot(type = "heatmap")

classification_knn |> get_pr()
```

***

With the method of K-nearest neighbours we can take a deeper look into the specifics of the clustering algorithm to see what is going on. 




### AI or no AI

```{r}

compmus_cv <- compmus2025_filtered |> vfold_cv(5)
compmus2025_filtered |>
  ggplot(aes(x = valence, y = arousal, colour = ai, size = tempo)) +
  geom_point(alpha = 0.8) +
  scale_color_viridis_d() +
  labs(
    x = "Valence",
    y = "Arousal",
    size = "Tempo",
    colour = "AI"
  )
```

### spectral & energy novelty

```{r, energy_novelty}
  data1 <-compmus_energy_novelty("features/popke-s-1.json") 
  fig1 <- (ggplot(data1, aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Energy Novelty song1", x = "Time (s)", y = "Energy Novelty"))


  data2 <- compmus_energy_novelty("features/popke-s-2.json")
  fig2 <- (ggplot(data2, aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Energy Novelty song 2",x = "Time (s)", y = "Energy Novelty"))

  data3 <- compmus_spectral_novelty("features/popke-s-1.json")
  fig3 <- (ggplot(data3, aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Spectral Novelty song 1", x = "Time (s)", y = "Spectral Novelty")) 
  
  data4 <- compmus_spectral_novelty("features/popke-s-2.json")
  fig4 <- (ggplot(data4, aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Spectral Novelty song 2", x = "Time (s)", y = "Spectral Novelty"))

  
  fig1 / fig2 / fig3 / fig4

```

***

information. more information

### Tempograms

```{r, tempograms}
  data1 <- compmus_tempogram("features/popke-s-1.json", window_size = 4, hop_size = 2, cyclic = FALSE)
  fig1 <- ggplot(data1, aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

  data2 <- compmus_tempogram("features/popke-s-2.json", window_size = 4, hop_size = 2, cyclic = FALSE)
  fig2 <- ggplot(data2, aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
  
  fig <- subplot(fig1, fig2, nrows=2)%>%
  layout(
    title = "tempograms",
    xaxis = list(title = "Time"),
    yaxis = list(title = "Temo (BPM)")
  )
  fig
```


***
Here we see the tempo gram plots for both songs. I first started out with a window size of 32 and hop size of 4. Where-after i got only really tight lines in my tempo grams. I adjusted the window size first to 8 and kept the hop size the same. The fluctuations in the tempo gram became already clearer. With lowering the window size you can clearly see more fluctuations and details in the tempo gram. We can see a clear tempo at ±135 BPM, which is also the tempo of the song, then we see 3 subharmonics at ±275, ±405, ±524. These are all doublings of the base tempo and therefore picked up in the tempograph. 


### cyclic tempograms

```{r, cyclic tempograms}
  data1 <- compmus_tempogram("features/popke-s-1.json", window_size = 4, hop_size = 2, cyclic = TRUE)
  fig1 <- ggplot(data1, aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()

  data2 <- compmus_tempogram("features/popke-s-2.json", window_size = 4, hop_size = 2, cyclic = TRUE)
  fig2 <- ggplot(data2, aes(x = time, y = bpm, fill = power)) +
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = "Tempo (BPM)") +
  theme_classic()
  
  fig <- subplot(fig1, fig2, nrows=2)%>%
  layout(
    title = "tempograms",
    xaxis = list(title = "Time"),
    yaxis = list(title = "Tempo (BPM)")
  )
  fig
```

***

In the cyclic tempo gram, we have a more visible sight of the actual tempo of the songs. Which is arond 137 for both of them. 


### keychrams and Chordchrams 

```{r, Keycharams & Choardchrams}

  data1 <- compmus_chroma("features/popke-s-1.json", norm = "identity")%>%
  compmus_match_pitch_templates(
    key_templates,         
    norm = "identity",       
    distance = "cosine")
  
  fig1 <- ggplot(data1, aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +
  labs(title = "keygram song 1", y = "Keys", fill = NULL) + theme_classic() + theme(plot.title = element_text(size=8), axis.title.y = element_text(size=8))

  data2 <- compmus_chroma("features/popke-s-1.json", norm = "identity")%>%
    compmus_match_pitch_templates(
    chord_templates,         
    norm = "identity",       
    distance = "cosine"   
  ) 
  
  fig2 <- ggplot(data2, aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +           
  labs(title = "chordgram song 1", y = "Chords", fill = NULL) + theme_classic() + theme(plot.title = element_text(size=8),  axis.title.y = element_text(size=8))

    data3 <- compmus_chroma("features/popke-s-2.json", norm = "identity")%>%
  compmus_match_pitch_templates(
    key_templates,         
    norm = "identity",       
    distance = "cosine")
  
  fig3 <- ggplot(data3, aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               
  labs(title = "keygram song 2", y = "Keys", fill = NULL) + theme_classic() + theme(plot.title = element_text(size=8), axis.title.y = element_text(size=8))

  data4 <- compmus_chroma("features/popke-s-2.json", norm = "identity")%>%
    compmus_match_pitch_templates(
    chord_templates,         
    norm = "identity",       
    distance = "cosine"   
  ) 
  fig4 <- ggplot(data4, aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               
  labs(title = "chordgram song 2", x = "Time (s)", y = "Chords", fill = NULL) + theme_classic() + theme(plot.title = element_text(size=8), axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
  
fig1 / fig2 / fig3 / fig4 
  
```

***
Here we see a representation of key grams and chord grams for both songs. They give us a view in the melody played troughout the songs. 

### Chromagrams song 1 & 2

```{r, Chromagrams}
  data1 <- compmus_chroma("features/popke-s-1.json", norm = "identity")      
  fig1 <- ggplot(data1, aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()  

  data2 <- compmus_chroma("features/popke-s-2.json", norm = "identity")
  fig2 <- ggplot(data2, aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +              
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()  
  
  fig <- subplot(fig1, fig2)
  fig
```

***

A chromagram is a time versus pitch class matrix, which is often used in Music Information Retrieval (MIR). It shows how much energy there is present for each of the 12 pitch classes. While a spectogram shows all frequencies, a chromagram visualises the frequencies for each octave in a single representation. 
When looking at the left chromagram, we can see that it appears a bit denser with a few more bright patches which suggest the use of louder notes or more harmonic complexity. The right chromagram has a more evenly divided distribution. We can see a lot of energy in the tonal use at the E, F and G chords. 
This could indicate that song 2 is played in an F major or D minor scale whereas song 1 could be played in A minor or G major. 


### cepstograms song 1 and 2 

```{r, Cepstograms}
  data1 <- compmus_mfccs("features/popke-s-1.json", norm = "identity")
  fig1 <- ggplot(data1, aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +              
  labs(x = "Time (s)", y = "Coefficient Number song 1", fill = NULL) +
  theme_classic()                                

  data2 <- compmus_mfccs("features/popke-s-2.json", norm = "identity")
  fig2 <- ggplot(data2, aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +              
  labs(x = "Time (s)", y = "Coefficient Number song 2", fill = NULL) +
  theme_classic()

 fig1 / fig2 

```

***
This image shows the two cepstograms for both songs, they visualize the evolution of Mel-Fequency Cepstral Coefficients (MFCCs) over time. MFCCs are used for speech and audio processing because they capture the spectral characteristics of timbre and phonetic features. 
Both cepstograms share a similar overall structure. The right one shows a more smoothed and uniform distribution, which indicates a more stable spectral envelope. Which fits with the song, where tones are very stable and there are not many fluctuations. 
The left cepstogram shows more variation in the horizontal band on the bottom, in the blue part. This indicates more energy in the lower cepstral coefficients. We can distinguish a difference in spectral energy in both songs. 


### Data applied on test visualisation

```{r,datavisualisation}
aisc2024 <- read_csv("compmus2025.csv")
aisc2024 |>                   # Start with the data
  ggplot(                     # Set up the plot.
    aes(
      x = tempo,
      y = arousal,
      size = instrumentalness,
      colour = danceability
    )
  ) +
  geom_point() +              # Scatter plot.
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
  geom_text(                  # Add text labels from above.
    x = 121,
    y = 4.91,
    label = "Onda Corta - Sud America",
    size = 3,                 # Override size (not loudness here).
    hjust = "left",           # Align left side of label with the point.
    vjust = "center",         # Align vertical centre of label with the point.
    angle = 30                # Rotate the text label
  ) +
  scale_x_continuous(         # Fine-tune the x axis.
    limits = c(50, 200),
    breaks = c(50, 100, 150, 200), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(1, 9),
    breaks = c(1, 5, 9),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # Use the popular viridis colour palette.
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Tempo",
    y = "Arousal",
    colour = "Danceability"
  )
ggplotly()
```


### histogram tempo

```{r, histogram data}

cm2025 <- read_csv("compmus2025.csv")
  ggplot(cm2025, aes(x= tempo)) + geom_histogram()
  ggplotly()
```

***

This is a histogram that gives insight to the tempo variation of our class corpus. The median is at 85 and the mean is around 100. 
I think the analysis isn't really specific, for both my own songs the tempo was 30 bpm lower than i saw in the file. I think the data is not that reliable. 


### correlation tempo and danceability

```{r, correlation tempo/danceability}
cm2025 <- read_csv("compmus2025.csv")
ggplot(cm2025, aes(x = tempo, y = danceability)) +
  geom_point(aes(color = engagingness), alpha = 0.7) +
  theme_minimal() +
  labs(title = "Tempo vs. Danceability", x = "Tempo", y = "Danceability") +
  scale_color_gradient(low = "purple", high = "orange")
ggplotly()
```




