---
title: "compmus"
author: "popke"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(plotly)
source("compmus.R")
library(rjson)
```

### Welcome

This dashboard will be a demonstration of correlation between various musical terms, were we will look for correlations between music in various ways. By using the music metrics of the music corpus of the computational musicology class of 2025 we will try to get a better understanding of music and sound. 

I've chosen my tracks because one of them is made by artist I work with and am a huge fan of and the other is made by a friend who also produces music. The tracks are quiet different, which I like because I hope I can compare them in interesting ways. I think my tracks match the class corpus, I came across a lot of instrumental music with a lot melodies in it. I think it's an interesting mix of music. 

### spectral & energy novelty

```{r, energy_novelty, cache=TRUE}
  data1 <-compmus_energy_novelty("features/popke-s-1.json") 
  fig1 <- ggplot(data1, aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Energy Novelty")


  data2 <- compmus_energy_novelty("features/popke-s-2.json")
  fig2 <- ggplot(data2, aes(t, novelty)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Time (s)", y = "Energy Novelty")
  
  fig <- subplot(fig1, fig2) %>%
  layout(
    title = "energy novelty",
    xaxis = list(title = "Time"),
    yaxis = list(title = "Energy Novelty")
  )
  fig
```

```{r, sepctral_novelty, cache=TRUE}

```


### keychrams and Chordchrams song 1

```{r, cache=TRUE}
#      C     C#    D     Eb    E     F     F#    G     Ab    A     Bb    B
major_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    0,    0)
minor_chord <-
  c(   1,    0,    0,    1,    0,    0,    0,    1,    0,    0,    0,    0)
seventh_chord <-
  c(   1,    0,    0,    0,    1,    0,    0,    1,    0,    0,    1,    0)

major_key <-
  c(6.35, 2.23, 3.48, 2.33, 4.38, 4.09, 2.52, 5.19, 2.39, 3.66, 2.29, 2.88)
minor_key <-
  c(6.33, 2.68, 3.52, 5.38, 2.60, 3.53, 2.54, 4.75, 3.98, 2.69, 3.34, 3.17)

chord_templates <-
  tribble(
    ~name, ~template,
    "Gb:7", circshift(seventh_chord, 6),
    "Gb:maj", circshift(major_chord, 6),
    "Bb:min", circshift(minor_chord, 10),
    "Db:maj", circshift(major_chord, 1),
    "F:min", circshift(minor_chord, 5),
    "Ab:7", circshift(seventh_chord, 8),
    "Ab:maj", circshift(major_chord, 8),
    "C:min", circshift(minor_chord, 0),
    "Eb:7", circshift(seventh_chord, 3),
    "Eb:maj", circshift(major_chord, 3),
    "G:min", circshift(minor_chord, 7),
    "Bb:7", circshift(seventh_chord, 10),
    "Bb:maj", circshift(major_chord, 10),
    "D:min", circshift(minor_chord, 2),
    "F:7", circshift(seventh_chord, 5),
    "F:maj", circshift(major_chord, 5),
    "A:min", circshift(minor_chord, 9),
    "C:7", circshift(seventh_chord, 0),
    "C:maj", circshift(major_chord, 0),
    "E:min", circshift(minor_chord, 4),
    "G:7", circshift(seventh_chord, 7),
    "G:maj", circshift(major_chord, 7),
    "B:min", circshift(minor_chord, 11),
    "D:7", circshift(seventh_chord, 2),
    "D:maj", circshift(major_chord, 2),
    "F#:min", circshift(minor_chord, 6),
    "A:7", circshift(seventh_chord, 9),
    "A:maj", circshift(major_chord, 9),
    "C#:min", circshift(minor_chord, 1),
    "E:7", circshift(seventh_chord, 4),
    "E:maj", circshift(major_chord, 4),
    "G#:min", circshift(minor_chord, 8),
    "B:7", circshift(seventh_chord, 11),
    "B:maj", circshift(major_chord, 11),
    "D#:min", circshift(minor_chord, 3)
  )

key_templates <-
  tribble(
    ~name, ~template,
    "Gb:maj", circshift(major_key, 6),
    "Bb:min", circshift(minor_key, 10),
    "Db:maj", circshift(major_key, 1),
    "F:min", circshift(minor_key, 5),
    "Ab:maj", circshift(major_key, 8),
    "C:min", circshift(minor_key, 0),
    "Eb:maj", circshift(major_key, 3),
    "G:min", circshift(minor_key, 7),
    "Bb:maj", circshift(major_key, 10),
    "D:min", circshift(minor_key, 2),
    "F:maj", circshift(major_key, 5),
    "A:min", circshift(minor_key, 9),
    "C:maj", circshift(major_key, 0),
    "E:min", circshift(minor_key, 4),
    "G:maj", circshift(major_key, 7),
    "B:min", circshift(minor_key, 11),
    "D:maj", circshift(major_key, 2),
    "F#:min", circshift(minor_key, 6),
    "A:maj", circshift(major_key, 9),
    "C#:min", circshift(minor_key, 1),
    "E:maj", circshift(major_key, 4),
    "G#:min", circshift(minor_key, 8),
    "B:maj", circshift(major_key, 11),
    "D#:min", circshift(minor_key, 3)
  )

 
  data1 <- compmus_chroma("features/popke-s-1.json", norm = "identity")%>%
  compmus_match_pitch_templates(
    key_templates,         
    norm = "identity",       
    distance = "cosine")
  
  fig1 <- ggplot(data1, aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic() 

  data2 <- compmus_chroma("features/popke-s-1.json", norm = "identity")%>%
    compmus_match_pitch_templates(
    chord_templates,         
    norm = "identity",       
    distance = "cosine"   
  ) 
  fig2 <- ggplot(data2, aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()  
  
    data3 <- compmus_chroma("features/popke-s-2.json", norm = "identity")%>%
  compmus_match_pitch_templates(
    key_templates,         
    norm = "identity",       
    distance = "cosine")
  
  fig3 <- ggplot(data3, aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic() 

  data4 <- compmus_chroma("features/popke-s-2.json", norm = "identity")%>%
    compmus_match_pitch_templates(
    chord_templates,         
    norm = "identity",       
    distance = "cosine"   
  ) 
  fig4 <- ggplot(data4, aes(x = time, y = name, fill = d)) + 
  geom_raster() +
  scale_fill_viridis_c(guide = "none") +               
  labs(x = "Time (s)", y = "Template", fill = NULL) +
  theme_classic()  
  
  fig <- subplot(fig1, fig2, fig3, fig4, nrows = 2) %>%
    layout(title = "Keygrams and Chord grams for song 1 & 2")
  fig
  
```


### Chromagrams song 1 & 2

```{r, cache=TRUE}
  data1 <- compmus_chroma("features/popke-s-1.json", norm = "identity")      
  fig1 <- ggplot(data1, aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()  

  data2 <- compmus_chroma("features/popke-s-2.json", norm = "identity")
  fig2 <- ggplot(data2, aes(x = time, y = pc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:11,
    minor_breaks = NULL,
    labels = c(
                "C", "C#|Db", "D", "D#|Eb",
                "E", "F", "F#|Gb", "G",
                "G#|Ab", "A", "A#|Bb", "B"
              )
  ) +
  scale_fill_viridis_c(guide = "none") +              
  labs(x = "Time (s)", y = NULL, fill = NULL) +
  theme_classic()  
  
  fig <- subplot(fig1, fig2)
  fig
```


### cepstograms song 1 and 2 

```{r, cache=TRUE}
  data1 <- compmus_mfccs("features/popke-s-1.json", norm = "identity")
  fig1 <- ggplot(data1, aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +              
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic()                                 

  data2 <- compmus_mfccs("features/popke-s-2.json", norm = "identity")
  fig2 <- ggplot(data2, aes(x = time, y = mfcc, fill = value)) + 
  geom_raster() +
  scale_y_continuous(
    breaks = 0:12,
    minor_breaks = NULL,
  ) +
  scale_fill_viridis_c(guide = "none") +              
  labs(x = "Time (s)", y = "Coefficient Number", fill = NULL) +
  theme_classic() 

fig <- subplot(fig1, fig2) 
fig
```


### Data applied on test visualisation

```{r, cache=TRUE}
aisc2024 <- read_csv("compmus2025.csv")
aisc2024 |>                   # Start with the data
  ggplot(                     # Set up the plot.
    aes(
      x = tempo,
      y = arousal,
      size = instrumentalness,
      colour = danceability
    )
  ) +
  geom_point() +              # Scatter plot.
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
  geom_text(                  # Add text labels from above.
    x = 121,
    y = 4.91,
    label = "Onda Corta - Sud America",
    size = 3,                 # Override size (not loudness here).
    hjust = "left",           # Align left side of label with the point.
    vjust = "center",         # Align vertical centre of label with the point.
    angle = 30                # Rotate the text label
  ) +
  scale_x_continuous(         # Fine-tune the x axis.
    limits = c(50, 200),
    breaks = c(50, 100, 150, 200), # Specify grid lines
    minor_breaks = NULL       # Remove 'minor' grid lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(1, 9),
    breaks = c(1, 5, 9),
    minor_breaks = NULL
  ) +
  scale_colour_viridis_c() +  # Use the popular viridis colour palette.
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud..
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "Tempo",
    y = "Arousal",
    colour = "Danceability"
  )
ggplotly()
```


### histogram tempo

```{r, cache=TRUE}

cm2025 <- read_csv("compmus2025.csv")
  ggplot(cm2025, aes(x= tempo)) + geom_histogram()
  ggplotly()
```


### correlation tempo and danceability

```{r, cache=TRUE}
cm2025 <- read_csv("compmus2025.csv")
ggplot(cm2025, aes(x = tempo, y = danceability)) +
  geom_point(aes(color = engagingness), alpha = 0.7) +
  theme_minimal() +
  labs(title = "Tempo vs. Danceability", x = "Tempo", y = "Danceability") +
  scale_color_gradient(low = "purple", high = "orange")
ggplotly()
```



Column {data-width=350}
-----------------------------------------------------------------------

### Toughts 

So far we can see there is a correlation between tempo and danceablility. This is still a work in progress page. 

